Privacy-Preserving Localization Using Private Linear-Combination Aggregation

### 2021-06-01 main.tex
Big progress on the introduction

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

@@

\begin{document}
%
% paper title
\title{Privacy-Preserving Localization Using Private Linear-Combination Aggregation}
%
%
% author names and IEEE memberships
\author{Marko~Ristic$^{1}$,
        Benjamin~Noack$^{2}$,~\IEEEmembership{Member,~IEEE,}
        and~Uwe~D.~Hanebeck$^{1}$,~\IEEEmembership{Fellow,~IEEE}% <-this % stops a space
\thanks{$^{1}$Marko Ristic and Uwe D. Hanebeck are with the Intelligent Sensor-Actuator-Systems Laboratory (ISAS), Institute for Anthropomatics, Karlsruhe Institute of Technology (KIT), Germany (e-mail: \{marko.ristic, uwe.hanebeck\}@kit.edu).}% <-this % stops a space
\thanks{$^{2}$Benjamin Noack is with the Autonomous Multisensor Systems Group (AMS), Institute for Intelligent Cooperating Systems (ICS), Otto von Guericke University Magdeburg (OVGU), Germany (e-mail: benjamin.noack@ovgu.de).}}% <-this % stops a space
% \thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% The paper headers
% \markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
% {Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}

% If you want to put a publisher's ID mark on the page you can do it like this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localization methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localization protocols, thus not preserving participants’ data privacy. Existing approaches that do provide such guarantees fail to address sensor and navigator privacy in common non-linear measurement model applications and therefore forfeit broad applicability. We define a suitable notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Kalman Filter with range-sensor measurements. Our method keeps navigator location, sensors' locations, and sensors' measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localization filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

@@

\section{Introduction} \label{sec:introduction}

\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}.

Range-based localisation methods, with measurements such as signal strength \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range \cite{beutlerNewNonlinearFiltering2004} or radar, have found particularly large application due to the prevalence of suitable sensors.

most rely on centralised

with the rise of distributed cloud

%starts to broad, skip to range-based?
%As advancements in portable computing and sensor capabilities have developed, a wide variety of localisation methods have been proposed, including those using signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range measurements \cite{beutlerNewNonlinearFiltering2004}, magnetic field measurements \cite{sieblerLocalizationMagneticField2020} and time-difference-of-arrival multilateration methods \cite{liContributedReviewSourcelocalization2016} to localisation agents.

%Due to the prevalence of suitable sensors such as radar or signal strength measurement, range-based localisation methods have found particularly large application.

Localisation methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localization methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004}, and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localization methods, including the use of radar measurements, GPS clocks, and signal strength measurements, have found particularly large applications due to their robustness and accuracy.

These methods use distance measurements to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location, and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}.

For example, an aircraft localization scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighboring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations.

All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows the private aggregation of weighted sums, before developing a computationally plausible localization filter in the presence of range-only sensors using this scheme, such that sensor and navigator privacy is preserved.

% LITERATURE OVERVIEW

In this section, we will discuss methods for providing security guarantees in signal processing tasks, and their applications to localization and estimation problems.

% ENCRYPTION SCHEMES IN SIGNAL PROCESSING

Of the methods used to preserve privacy during distributed signal processing, it is common for the information exchanging portion of the algorithm to be reduced to a form of aggregation of involved parties' data. The most common methods for ensuring privacy during aggregation required for signal processing tasks have been described below.
    \item[Differential Privacy] These techniques formalize preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation. To handle the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs \cite{hanPrivacyControlDynamical2018,dworkDifferentialPrivacySurvey2008,andresGeoindistinguishabilityDifferentialPrivacy2013}. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localization.
    \item[Aggregator Oblivious Encryption] This method of encryption formalizes the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their aggregation (typically sum) can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.
In addition to the private aggregation of data, it may also be the case that private data needs to be modified by an external party before being returned, as with privacy-preserving cloud computing. These methods have been broadly grouped into two domains and described below.
    \item[Fully Homomorphic Encryption] These schemes allow all algebraic computations to be performed on encrypted data without the computing party learning any information about inputs, intermediate values or results \cite{gentryFullyHomomorphicEncryption2009,stehleFasterFullyHomomorphic2010}. Although theoretically ideal for many data-sensitive tasks, current implementations are still computationally impractical for real-time or large scale tasks \cite{acarSurveyHomomorphicEncryption2018}.
    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes; these schemes provide only a subset of (typically one) algebraic operations on encrypted data \cite{elgamalPublicKeyCryptosystem1985,paillierPublicKeyCryptosystemsBased1999,bonehEvaluating2DNFFormulas2005}. Reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication \cite{kogisoCyberSecurityEnhancementNetworked2015}, set intersection computation \cite{kerschbaumOutsourcedPrivateSet2012} and control input aggregation \cite{alexandruEncryptedCooperativeControl2019}. However, as with Aggregator Oblivious Encryption, these works are relatively restricted in application due to the limited operations provided.
Lastly, recent developments in encryption schemes have presented some novel methods also suitable for privacy-preserving computation in distributed environments.
    \item[Multi-Client Functional Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. They are suitable for data-sensitive distributed processing as encryptions must be combined before a result is obtained, and can be considered a generalization of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.
In the context of privacy-preserving localization and state estimation, some of the aforementioned schemes have found applications and some specialized encryption methods have been proposed. These are summarized in the next section.

% APPLICATIONS TO LOCALIZATION AND ESTIMATION

Localization methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide an accurate localization filter in the presence of range-only sensors while providing uncertainties associated with estimates.

Relevant estimation methods include both Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and time-independent methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute time-independent location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided, making them unsuitable for our desired solution. Bayesian approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements.

Encrypted model-predictive-control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. In \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, the novel security notions of private weighted sum aggregation with centralized or hidden weights (pWSAc and pWSAh, respectively) are proposed. The schemes capture a weighted aggregation operation suitable for providing privacy in some control environments and are similar in principle to the aggregation decomposition proposed in this work. The requirements of our aggregation operation differ, however, in that weights are generalized to arbitrary length linear-combinations, which must be consistent across all participants, and can vary over time. We also require a formal security proof capturing these weight requirements as well as the explicit capabilities of participants. We define our alternative security notion in Section .

% Used to be in problem formulation but better point out the differences in our contribution to pWSAc here and not mention it again
Although similar to the pWSAc communication protocol in \cite{alexandruPrivateWeightedSum2020}, we formalise the assumptions

% Was in problem formulation but all motivations should be in the introduction instead
We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

Along with encryption schemes and their application to localization-relevant algorithms, encryption schemes with homomorphic properties also require suitable encoding when used with real numbers. As is the case with range sensors, real-valued measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{googleEncryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We use a modified version of \cite{farokhiSecurePrivateControl2017} due to its simplicity and applicability, and introduce the scheme details in Section .

While some aspects of privacy-preserving localization are solved in the above methods and encryption schemes, they fail to provide a localization solution in our context. In the next section, we will formalize our problem and its accompanying security requirements.

>>>>

@@

% Start
\begin{document}

% paper title
\title{Privacy-Preserving Localization Using Private Linear-Combination Aggregation}

% author names and IEEE memberships
\author{Marko~Ristic$^{1}$,
        Benjamin~Noack$^{2}$,~\IEEEmembership{Member,~IEEE,}
        and~Uwe~D.~Hanebeck$^{1}$,~\IEEEmembership{Fellow,~IEEE}% <-this % stops a space
\thanks{$^{1}$Marko Ristic and Uwe D. Hanebeck are with the Intelligent Sensor-Actuator-Systems Laboratory (ISAS), Institute for Anthropomatics, Karlsruhe Institute of Technology (KIT), Germany (e-mail: \{marko.ristic, uwe.hanebeck\}@kit.edu).}% <-this % stops a space
\thanks{$^{2}$Benjamin Noack is with the Autonomous Multisensor Systems Group (AMS), Institute for Intelligent Cooperating Systems (ICS), Otto von Guericke University Magdeburg (OVGU), Germany (e-mail: benjamin.noack@ovgu.de).}}% <-this % stops a space

@@

\begin{abstract}
Distributed state estimation and localization methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localization protocols, thus not preserving participants’ data privacy. Existing approaches that do provide such guarantees fail to address sensor and navigator privacy in common non-linear measurement model applications and therefore forfeit broad applicability. We define a suitable notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Kalman Filter with range-sensor measurements. Our method keeps navigator location, sensors' locations, and sensors' measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localization filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

@@

\section{Introduction} \label{sec:introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,ligginsDistributedDataFusion2012,liContributedReviewSourcelocalization2016} and have characterised many advancements of Kalman and Bayesian estimation theory \cite{mutambaraDecentralizedEstimationControl1998}. In particular, range-based localisation methods, including signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic ranges \cite{beutlerNewNonlinearFiltering2004} and ultra wideband ranges \cite{geziciLocalizationUltraWidebandRadios2005}, have found large application due to the prevalence of suitable sensors. In most cases, these localisation methods require the gathering of measurements centrally, where an estimate of location can be computed. With recent developments in distributed and cloud computing, uses of wireless and public communication channels for data transfer have becoming widespread, and the additional requirements of data privacy and state secrecy have become particularly relevant \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}.

Typical cryptographic secrecy involves hiding all transferred data such that untrusted participants in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in such signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, respectively, are introduced as means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work considers range-based localisation with formal cryptographic requirements. We present a novel scheme satisfying the requirements before deriving an estimation filter which uses the scheme to allow localisation without learning sensor measurements, measurement errors or locations, while sensors learn nothing other than their own measurements.

based on the extended kalman filter, with no system model or sensor layout assumptions.

Our contribution considers the context of range-based localisation, and consists of a formal cryptographic definition for a novel linear cmobination aggregation scheme

Our contribution is in two parts

Although similar to the pWSAc communication protocol in \cite{alexandruPrivateWeightedSum2020}, we formalise the assumptions

Lastly, we motivate the scenario with some examples

% old one

% For example, an aircraft localization scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighboring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations.

% All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows the private aggregation of weighted sums, before developing a computationally plausible localization filter in the presence of range-only sensors using this scheme, such that sensor and navigator privacy is preserved.

% % ENCRYPTION SCHEMES IN SIGNAL PROCESSING

% Of the methods used to preserve privacy during distributed signal processing, it is common for the information exchanging portion of the algorithm to be reduced to a form of aggregation of involved parties' data. The most common methods for ensuring privacy during aggregation required for signal processing tasks have been described below.
% \begin{LaTeXdescription}
%     \item[Differential Privacy] These techniques formalize preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation. To handle the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs \cite{hanPrivacyControlDynamical2018,dworkDifferentialPrivacySurvey2008,andresGeoindistinguishabilityDifferentialPrivacy2013}. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localization.
%     \item[Aggregator Oblivious Encryption] This method of encryption formalizes the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their aggregation (typically sum) can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.
% \end{LaTeXdescription}
% In addition to the private aggregation of data, it may also be the case that private data needs to be modified by an external party before being returned, as with privacy-preserving cloud computing. These methods have been broadly grouped into two domains and described below.
% \begin{LaTeXdescription}
%     \item[Fully Homomorphic Encryption] These schemes allow all algebraic computations to be performed on encrypted data without the computing party learning any information about inputs, intermediate values or results \cite{gentryFullyHomomorphicEncryption2009,stehleFasterFullyHomomorphic2010}. Although theoretically ideal for many data-sensitive tasks, current implementations are still computationally impractical for real-time or large scale tasks \cite{acarSurveyHomomorphicEncryption2018}.
%     \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes; these schemes provide only a subset of (typically one) algebraic operations on encrypted data \cite{elgamalPublicKeyCryptosystem1985,paillierPublicKeyCryptosystemsBased1999,bonehEvaluating2DNFFormulas2005}. Reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication \cite{kogisoCyberSecurityEnhancementNetworked2015}, set intersection computation \cite{kerschbaumOutsourcedPrivateSet2012} and control input aggregation \cite{alexandruEncryptedCooperativeControl2019}. However, as with Aggregator Oblivious Encryption, these works are relatively restricted in application due to the limited operations provided.
% \end{LaTeXdescription}
% Lastly, recent developments in encryption schemes have presented some novel methods also suitable for privacy-preserving computation in distributed environments.
% \begin{LaTeXdescription}
%     \item[Multi-Client Functional Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. They are suitable for data-sensitive distributed processing as encryptions must be combined before a result is obtained, and can be considered a generalization of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.
% \end{LaTeXdescription}
% In the context of privacy-preserving localization and state estimation, some of the aforementioned schemes have found applications and some specialized encryption methods have been proposed. These are summarized in the next section.

% % APPLICATIONS TO LOCALIZATION AND ESTIMATION

% Localization methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide an accurate localization filter in the presence of range-only sensors while providing uncertainties associated with estimates.

% Relevant estimation methods include both Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and time-independent methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute time-independent location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided, making them unsuitable for our desired solution. Bayesian approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements.

% Encrypted model-predictive-control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. In \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, the novel security notions of private weighted sum aggregation with centralized or hidden weights (pWSAc and pWSAh, respectively) are proposed. The schemes capture a weighted aggregation operation suitable for providing privacy in some control environments and are similar in principle to the aggregation decomposition proposed in this work. The requirements of our aggregation operation differ, however, in that weights are generalized to arbitrary length linear-combinations, which must be consistent across all participants, and can vary over time. We also require a formal security proof capturing these weight requirements as well as the explicit capabilities of participants. We define our alternative security notion in Section .

% % Used to be in problem formulation but better point out the differences in our contribution to pWSAc here and not mention it again
% Although similar to the pWSAc communication protocol in \cite{alexandruPrivateWeightedSum2020}, we formalise the assumptions

% % Was in problem formulation but all motivations should be in the introduction instead
% We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

% Along with encryption schemes and their application to localization-relevant algorithms, encryption schemes with homomorphic properties also require suitable encoding when used with real numbers. As is the case with range sensors, real-valued measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{googleEncryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We use a modified version of \cite{farokhiSecurePrivateControl2017} due to its simplicity and applicability, and introduce the scheme details in Section .

% While some aspects of privacy-preserving localization are solved in the above methods and encryption schemes, they fail to provide a localization solution in our context. In the next section, we will formalize our problem and its accompanying security requirements.
